Agenda diversity and coverage bias: A quantitative approach to the agenda-setting theory

### 2018-07-30 plos_latex_template.tex
Primera version pasada a formato plos one

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Agenda diversity and coverage bias: A quantitative approach to the agenda-setting theory}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
   email={spinto@df.uba.ar}   % email address
]{\inits{SP}\fnm{Sebasti\'an} \snm{Pinto}}
\author[
   addressref={aff3},                   % id's of addresses, e.g. {aff1,aff2}
]{\inits{FA}\fnm{Federico} \snm{Albanese}}
\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
]{\inits{COD}\fnm{Claudio O} \snm{Dorso}}
\author[
   addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
   email={balen@df.uba.ar}   % email address
]{\inits{PB}\fnm{Pablo} \snm{Balenzuela}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgname{Departamento de F\'isica, Facultad de Ciencias Exactas y Naturales, Universidad de Buenos Aires}, % university, etc
  \street{Av. Cantilo s/n, Pabell\'on 1, Ciudad Universitaria},                     %
  \postcode{1428},                                % post or zip code
  \city{Buenos Aires},                              % city
  \cny{AR}                                    % country
}
\address[id=aff2]{%                           % unique id
  \orgname{Instituto de F\'isica de Buenos Aires (IFIBA), CONICET}, % university, etc
  \street{Av. Cantilo s/n, Pabell\'on 1, Ciudad Universitaria},                     %
  \postcode{1428},                                % post or zip code
  \city{Buenos Aires},                              % city
  \cny{AR}                                    % country
}
\address[id=aff3]{%                           % unique id
  \orgname{Instituto de Investigaci\'on en Ciencias de la Computaci\'on (ICC), CONICET}, % university, etc
  \street{Av. Cantilo s/n, Pabell\'on 1, Ciudad Universitaria},                     %
  \postcode{1428},                                % post or zip code
  \city{Buenos Aires},                              % city
  \cny{AR}                                    % country
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract

The mass media play a fundamental role in the formation of public opinion, either by defining the topics of discussion or by making a degree of emphasis on certain issues. Directly or indirectly, people get informed by consuming news from the media. But which is the dynamics of the agenda and how the people become interested in the different topics of the agenda? The Agenda Setting theory provides a conceptual framework in order to understand the role played by the mass media in public opinion formation, but the previous questions can not be answered without proper quantitative measures of agenda's dynamics and public attention. In this work we study the agenda of Argentinian newspapers in comparison with public's interests through a quantitative approach by performing topic detection over the news, identifying the main topics covered and their evolution over time. We measure Agenda's diversity as a function of time using Shannon's entropy and difference between Agendas using Jensen-Shannon's distance. We found that the Public Agenda is less diverse than the Media Agenda, and we are also capable to detect periods of time where coverage of certain issues are biased (coverage bias).\\

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{agenda-setting}
\kwd{agenda diversity}
\kwd{coverage bias}
\kwd{topic modelling}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%

\section*{Introduction}

\subsection*{Agenda Setting Theory}

\par To understand the ecosystem of information flow and opinion formation is one of the goals in these days. In this framework, many ingredients can be part of different mechanism of social influence. A major role in this ecosystem is played by the Mass Media outlets, which used to be the source of information of many of us.  Informed people tends to interact with each other, either via personal interactions or through social networks. In this scenario, becomes essential to understand the role of Mass Media Influence in a given social group.
\par  This question has been raised years ago. In the famous study performed in Chapel Hill during the US presidential elections in 1968 \cite{mccombs1972agenda}, Maxwell McCombs and Donald Shaw found that those aspects of public affairs that are prominent in the news become prominent among the public.
This study is considered the founding of the agenda-setting theory, which focus in the influence of mass media in public opinion.  
From \cite{mccombs2014agenda}, \textit{``The media agenda is the pattern of news coverage over a period of days, weeks (...) for a set of issues or other topic. In other words, the media agenda is a systematic compilation of the issues or topics presented to the public that identifies the degree of emphasis on these topics''}. 
\par In its basic stage, the theory focus on the comparison between the topics coverage by the media and the public agenda, i.e. the topics that the public consider as priority.
It looks for answering the question if the media is able to set the public agenda, which would transform the media as an important actor in the formation of public opinion. 
This stage is sometimes called the ``first level agenda-setting''.  The very often quoted phrase of Bernard C. Cohen \textit{``The press may not be successful much of the time in telling people what to think, but it is stunningly successful in telling its readers what to think about.''} illustrates its object of study. With the emergency of the Internet, the end of agenda-setting were predicted due to the audience fragmentation onto multiple sources, which would virtually lead to a highly individualized agenda. However, it is based on two assumptions that are not necessarily true: the public spreads its attention in an homogeneous way across the multiple sources, and that the agendas of that sources are different \cite{mccombs2005look}.
\par From the Chapel Hill research, several directions of agenda-setting were established \cite{mccombs2005look}.
The ``second level agenda-setting'', sometimes called \textit{attribute agenda-setting}, studies the \textit{objects} (in a social psychology way, where an \textit{attitude object} designate a thing that an individual has an attitude or opinion about) present in the media agenda. When the media talks about an object some attributes are emphasized, and others not. 
This theory is linked with \textit{framing} \cite{guggenheim2015dynamics, tsur2015frame}. 
To frame is to \textit{select some aspects of a perceived reality and make them more salient in a communicating text, in such a way as to promote a particular problem definition, causal interpretation, moral evaluation and/or treatment recommendation} (Robert Entman) \cite{mccombs2005look}.
This stage of agenda-setting theory can be summary in the phrase \textit{``the media not only can be successful in telling us what to think about, they also can be successful in telling us how to think about it.''} 
Other interesting stage of agenda-setting concerns with the sources of media agenda, i.e. if the media set the public agenda, \textit{who sets the agenda media?} Within this framework, \textit{intermedia agenda-setting} observes the competition between different media and how they influence each other. The competition between mass media for the same audience can lead to a homogenization of the agendas \cite{vargo2017networks}, which is in the opposite direction of one of the assumptions that predict the end of agenda-setting, as was mentioned before.
\par Since the irruption of internet in our lives and the access to big data, there were many attempts to analyze from a quantitative point of view the theory of Agenda Setting based on  the original work of McCombs \cite{mccombs1972agenda}. Many of them follows what we called a ``static approach", i.e. by summarizing certain issues present in news media and looking for effects in the audience \cite{brians1996campaign, gerber2009does, coleman2007young}.
For instance, in Argentina, \cite{mitchelstein2016brecha} examines the difference between public and journalists preferences. 
Several works by E. Zunino, like \cite{zunino2010cobertura} and \cite{koziner2013cobertura}, studied the coverage given by the main newspapers of particular events which have happened during the former administration of Argentinian president \emph{Cristina Fern\'andez de Kirchner} (from 2007 to 2015), where the government started to confront with news organizations that were critical to its management as they were an opposition party \cite{mitchelstein2017information}.

@@

\par Most of the works cited above follow typical methods of social science research. However, a useful tool in the analysis of large corpus of documents, which is not widely used in the agenda setting theory, is unsupervised topic modeling. These methods are an alternative to the dictionary-based analysis, which is the most popular automated analysis approach \cite{guo2016big}, and allows to work with a corpus without a prior knowledge, letting the topics emerge from the data. Despite the popularity of these methods in other fields, they have been under-employed in the agenda setting framework.
Many works based on news corpus emphasize the performance of the topic model over a labeled corpus, focusing on the proper detection of the topics \cite{dai2010online, po2016topic, brun2000experiment}. The temporal profile of topics is usually embedded in the context of topic tracking \cite{hu2016news, li2017joint}, or in the recognition of emerging topics in real-time \cite{cataldi2010emerging} mostly applied to social media. 
\par A typical dynamical analysis of Agenda Setting is focused on a single issue. For instance, in \cite{soroka2017negativity} it is shown that the newspapers and Twitter have an opposite reaction to the changes of the unemployment rates. In \cite{guggenheim2015dynamics} the competition of frames about gun control is explored, and in \cite{ali2018measuring} the authors show how fluctuations of  twitter activity in different regions depend on the location of terrorist attacks.
A remarkable exception is found in  \cite{russell2014dynamics}, where they work with a set of predefined issues. In this work the question of \emph{causality} is also faced up. Their study shows that sometimes the traditional media set the agenda and sometimes the social media does. They show that social media is always more interested in social issues than the traditional one, and despite the existence of correlation, the social media agenda can not be seen as a \emph{slave} of the traditional media.
However, a drawback of working with predefined issues is that they usually reflects general themes, such as unemployment or crime, and the issues selection must be made by the researcher. 
\par In this work we propose a novel method in order to study the dynamics of Mass Media Agenda as a set of topic evolving in time and their correspondence with the Public and Social Network Agendas.  Our work intends to contribute another quantitative approach which complements the agenda-setting theory describe above. It mainly stands within the framework of first-level agenda setting, but issues about the comparison between media agendas and framing are also discussed.
Rather than focus on a single issue or a set of independent topics, we work with the agendas (the media and the public) as an object in their own, studying their evolution over time. 
On the other hand, we aim to take an insight about Media dynamics and Public response in order to create useful tools at the time of constructing mathematical models about the interaction between Media and Public, investigation that we started in \cite{pinto2016setting}.

\section*{Materials and Methods}
 \label{sec:MatMeth}

>>>>

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended 
% to minimize problems and delays during our production 
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % % 
%
% Once your paper is accepted for publication, 
% PLEASE REMOVE ALL TRACKED CHANGES in this file 
% and leave only the final text of your manuscript. 
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that 
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file. 
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission. 
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.  For example, x$^2$ is incorrect; this should be formatted as $x^2$ (or $\mathrm{x}^2$ if the romanized font is desired).
%
% Do not include text that is not math in the math environment. For example, CO2 should be written as CO\textsubscript{2} instead of CO$_2$.
%
% Please add line breaks to long display equations when possible in order to fit size of the column. 
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.  For example, change "[U(D,E,\gamma)]^2" to "{[U(D,E,\gamma)]}^2". 
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % % 
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

%% END MACROS SECTION

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Agenda diversity and coverage bias: A quantitative approach to the agenda-setting theory} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Sebasti\'an Pinto\textsuperscript{1,2,*},
Federico Albanese\textsuperscript{3},
Claudio O. Dorso\textsuperscript{1,2},
Pablo Balenzuela\textsuperscript{1,2}
\\
\bigskip
\textbf{1} Departamento de F\'isica, Facultad de Ciencias Exactas y Naturales, Universidad de Buenos Aires, Av. Cantilo s/n, Pabell\'on 1, Ciudad Universitaria, 1428, Buenos Aires, Argentina.
\\
\textbf{2} Instituto de F\'isica de Buenos Aires (IFIBA), CONICET, Av. Cantilo s/n, Pabell\'on 1, Ciudad Universitaria, 1428, Buenos Aires, Argentina.
\\
\textbf{3} Instituto de Investigaci\'on en Ciencias de la Computaci\'on (ICC), CONICET, Av. Cantilo s/n, Pabell\'on 1, Ciudad Universitaria, 1428, Buenos Aires, Argentina.
\\
\bigskip

% Insert additional author notes using the symbols described below. Insert symbol callouts after author names as necessary.
% 
% Remove or comment out the author notes below if they aren't used.
%
% Primary Equal Contribution Note
%\Yinyang These authors contributed equally to this work.

% Additional Equal Contribution Note
% Also use this double-dagger symbol for special authorship notes, such as senior authorship.
%\ddag These authors also contributed equally to this work.

% Current address notes
%\textcurrency Current Address: Dept/Program/Center, Institution Name, City, State, Country % change symbol to "\textcurrency a" if more than one current address note
% \textcurrency b Insert second current address 
% \textcurrency c Insert third current address

% Deceased author note
%\dag Deceased

% Group/Consortium Author Note
%\textpilcrow Membership list can be found in the Acknowledgments section.

% Use the asterisk to denote corresponding authorship and provide email address in note below.
* spinto@df.uba.ar

\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
The mass media play a fundamental role in the formation of public opinion, either by defining the topics of discussion or by making a degree of emphasis on certain issues. Directly or indirectly, people get informed by consuming news from the media. But which is the dynamics of the agenda and how the people become interested in the different topics of the agenda? The Agenda Setting theory provides a conceptual framework in order to understand the role played by the mass media in public opinion formation, but the previous questions can not be answered without proper quantitative measures of agenda's dynamics and public attention. In this work we study the agenda of Argentinian newspapers in comparison with public's interests through a quantitative approach by performing topic detection over the news, identifying the main topics covered and their evolution over time. We measure Agenda's diversity as a function of time using Shannon's entropy and difference between Agendas using Jensen-Shannon's distance. We found that the Public Agenda is less diverse than the Media Agenda, and we are also capable to detect periods of time where coverage of certain issues are biased (coverage bias).\\

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\section*{Author summary}

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\section*{Introduction}
\par To understand the ecosystem of information flow and opinion formation is one of the goals in these days. In this framework, many ingredients can be part of different mechanism of social influence. A major role in this ecosystem is played by the Mass Media outlets, which used to be the source of information of many of us.  Informed people tends to interact with each other, either via personal interactions or through social networks. In this scenario, becomes essential to understand the role of Mass Media Influence in a given social group.
\par  This question has been raised years ago. In the famous study performed in Chapel Hill during the US presidential elections in 1968 \cite{mccombs1972agenda}, Maxwell McCombs and Donald Shaw found that those aspects of public affairs that are prominent in the news become prominent among the public.
This study is considered the founding of the agenda-setting theory, which focus in the influence of mass media in public opinion.  
From \cite{mccombs2014agenda}, \textit{``The media agenda is the pattern of news coverage over a period of days, weeks (...) for a set of issues or other topic. In other words, the media agenda is a systematic compilation of the issues or topics presented to the public that identifies the degree of emphasis on these topics''}. 
\par In its basic stage, the theory focus on the comparison between the topics coverage by the media and the public agenda, i.e. the topics that the public consider as priority.
It looks for answering the question if the media is able to set the public agenda, which would transform the media as an important actor in the formation of public opinion. 
This stage is sometimes called the ``first level agenda-setting''.  The very often quoted phrase of Bernard C. Cohen \textit{``The press may not be successful much of the time in telling people what to think, but it is stunningly successful in telling its readers what to think about.''} illustrates its object of study. With the emergency of the Internet, the end of agenda-setting were predicted due to the audience fragmentation onto multiple sources, which would virtually lead to a highly individualized agenda. However, it is based on two assumptions that are not necessarily true: the public spreads its attention in an homogeneous way across the multiple sources, and that the agendas of that sources are different \cite{mccombs2005look}.

\par From the Chapel Hill research, several directions of agenda-setting were established \cite{mccombs2005look}.
The ``second level agenda-setting'', sometimes called \textit{attribute agenda-setting}, studies the \textit{objects} (in a social psychology way, where an \textit{attitude object} designate a thing that an individual has an attitude or opinion about) present in the media agenda. When the media talks about an object some attributes are emphasized, and others not. 
This theory is linked with \textit{framing} \cite{guggenheim2015dynamics, tsur2015frame}. 
To frame is to \textit{select some aspects of a perceived reality and make them more salient in a communicating text, in such a way as to promote a particular problem definition, causal interpretation, moral evaluation and/or treatment recommendation} (Robert Entman) \cite{mccombs2005look}.
This stage of agenda-setting theory can be summary in the phrase \textit{``the media not only can be successful in telling us what to think about, they also can be successful in telling us how to think about it.''} 
Other interesting stage of agenda-setting concerns with the sources of media agenda, i.e. if the media set the public agenda, \textit{who sets the agenda media?} Within this framework, \textit{intermedia agenda-setting} observes the competition between different media and how they influence each other. The competition between mass media for the same audience can lead to a homogenization of the agendas \cite{vargo2017networks}, which is in the opposite direction of one of the assumptions that predict the end of agenda-setting, as was mentioned before.

\par Since the irruption of internet in our lives and the access to big data, there were many attempts to analyze from a quantitative point of view the theory of Agenda Setting based on  the original work of McCombs \cite{mccombs1972agenda}. Many of them follows what we called a ``static approach", i.e. by summarizing certain issues present in news media and looking for effects in the audience \cite{brians1996campaign, gerber2009does, coleman2007young}.
For instance, in Argentina, \cite{mitchelstein2016brecha} examines the difference between public and journalists preferences. 
Several works by E. Zunino, like \cite{zunino2010cobertura} and \cite{koziner2013cobertura}, studied the coverage given by the main newspapers of particular events which have happened during the former administration of Argentinian president \emph{Cristina Fern\'andez de Kirchner} (from 2007 to 2015), where the government started to confront with news organizations that were critical to its management as they were an opposition party \cite{mitchelstein2017information}.

@@

\par Most of the works cited above follow typical methods of social science research. However, a useful tool in the analysis of large corpus of documents, which is not widely used in the agenda setting theory, is unsupervised topic modeling. These methods are an alternative to the dictionary-based analysis, which is the most popular automated analysis approach \cite{guo2016big}, and allows to work with a corpus without a prior knowledge, letting the topics emerge from the data. Despite the popularity of these methods in other fields, they have been under-employed in the agenda setting framework.
Many works based on news corpus emphasize the performance of the topic model over a labeled corpus, focusing on the proper detection of the topics \cite{dai2010online, po2016topic, brun2000experiment}. The temporal profile of topics is usually embedded in the context of topic tracking \cite{hu2016news, li2017joint}, or in the recognition of emerging topics in real-time \cite{cataldi2010emerging} mostly applied to social media.

\par A typical dynamical analysis of Agenda Setting is focused on a single issue. For instance, in \cite{soroka2017negativity} it is shown that the newspapers and Twitter have an opposite reaction to the changes of the unemployment rates. In \cite{guggenheim2015dynamics} the competition of frames about gun control is explored, and in \cite{ali2018measuring} the authors show how fluctuations of  twitter activity in different regions depend on the location of terrorist attacks.
A remarkable exception is found in  \cite{russell2014dynamics}, where they work with a set of predefined issues. In this work the question of \emph{causality} is also faced up. Their study shows that sometimes the traditional media set the agenda and sometimes the social media does. They show that social media is always more interested in social issues than the traditional one, and despite the existence of correlation, the social media agenda can not be seen as a \emph{slave} of the traditional media.
However, a drawback of working with predefined issues is that they usually reflects general themes, such as unemployment or crime, and the issues selection must be made by the researcher.

\par In this work we propose a novel method in order to study the dynamics of Mass Media Agenda as a set of topic evolving in time and their correspondence with the Public and Social Network Agendas.  Our work intends to contribute another quantitative approach which complements the agenda-setting theory describe above. It mainly stands within the framework of first-level agenda setting, but issues about the comparison between media agendas and framing are also discussed.
Rather than focus on a single issue or a set of independent topics, we work with the agendas (the media and the public) as an object in their own, studying their evolution over time. 
On the other hand, we aim to take an insight about Media dynamics and Public response in order to create useful tools at the time of constructing mathematical models about the interaction between Media and Public, investigation that we started in \cite{pinto2016setting}.

\section*{Materials and methods}

\subsection*{The Media Agenda}

In this work we analyze a three month period of the Argentinian Media's Agenda composed a corpus of news' articles that were published between July 31th and November 5th in the section \emph{Politics} of the electronic editions of the Argentinian newspapers \emph{Clar\'in}, \emph{La Naci\'on}, \emph{P\'agina12}, and the news portal \emph{Infobae}. The first two lead the sale of printed editions in \emph{Buenos Aires} city, but \emph{Clar\'in} reaches roughly two times the readers of \emph{La Naci\'on}, and ten times the readers of \emph{P\'agina 12} \cite{IVC}. On the other hand, \emph{Infobae} has the website with more visitors, above the websites of \emph{Clar\'in} and \emph{La Naci\'on} \cite{AlexaAR}.
The corpus analyzed is constituted by 2908 politics articles of \emph{Clar\'in}, 3565 of \emph{La Naci\'on}, 3324 of \emph{P\'agina 12}, and 2018 of \emph{Infobae}. Except \emph{P\'agina 12}, all articles were taken from the section \emph{Pol\'itica}  (Politics) of the respective news portal, while the articles which belong to \emph{P\'agina 12} were taken from the section \emph{El pa\'is} (The country).

\par In order to perform the analysis of the articles in the corpus, we describe them as numerical vectors through the \textit{term frequency - inverse document frequency (tf-idf)} representation \cite{xu2003document}. Given the set of terms contained in the corpus' words after removing non-informative ones, such as prepositions and conjunctions, the \textit{tf-idf} algorithm represents the \textit{i}-document as a vector $v_i = [x_{i1}, x_{i2}, ... , x_{it}]$, where the component $x_{ij}$ is computed by the equation \ref{ec:tfidf}, where $\textrm{tf}_{ij}$ is the number of times the \textit{j}-term appears in the \textit{i}-document, $d$ is the number of documents in the corpus, and $n_j$ is the number of documents where the \textit{j}-term appears. Each document's vector is normalized to unit Euclidean length. Once the documents' vector are constructed, we put them together in a document-term matrix (\emph{M}), which has dimensions of number of documents in the corpus ($d$) per number of terms selected ($t$).
